---
title: "Assignment 9"
author: "Erinda Budo"
date: "10/27/2019"
output:
  html_document:
    df_print: paged
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

**Assignment Overview**

The New York Times web site provides a rich set of APIs, as described here: https://developer.nytimes.com/apis(https://developer.nytimes.com/apis)

You’ll need to start by signing up for an API key.

Your task is to choose one of the New York Times APIs, construct an interface in R to read in the JSON data, and transform it into an R DataFrame.

**Load Package**

```{r,warning=FALSE,message=FALSE}
library(httr)
library(jsonlite)
library(knitr)
library(DT)
library(stringr)
library(dplyr)
```

**Read movie information**

```{r }
url<-("https://api.nytimes.com/svc/movies/v2/reviews/search.json?critics-pick=Y&order=by-title&api-key=eTlUJikeG81CLJ3wJYvBfxs6UAACelfi")
```


**Create a dataframe and display results**

```{r }
movie_reviews<- fromJSON(url) %>%
as.data.frame() %>%
select(5, 6, 7, 8)
names(movie_reviews) <- str_replace_all(names(movie_reviews), "results.", "")
datatable(movie_reviews)
```
**Fill null values in rating column**
```{r }
movie_reviews$mpaa_rating[movie_reviews$mpaa_rating == ''] <- 'Not Rated'
```
**Show bar chart of rating type**

```{r }
library(ggplot2)
movie_reviews %>%
   
    ggplot(aes(x=mpaa_rating, fill=mpaa_rating)) + geom_bar(position = "dodge") +# plot
    ggtitle("Movie Rating bar plot")
```


Top  reviews are for not rated movies


**Show bar chart of critics**

```{r }
movie_reviews %>%
   
    ggplot(aes(x=byline, fill=byline)) + geom_bar(position = "dodge") +# plot
    ggtitle("Critic bar plot")
```






Top reviews are done by :
Jeannette Catsoulis
Manohla Dargis
Stephen Holden