---
title: "Assignment R and SQL"
author: "Erinda Budo"
date: "9/7/2019"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

Introduction
The focus of this assignment is to create an R dataframe that shows rates of movies by gender of the population asked rating. Movies_db data will be sourced from a tb database in MySQL and combined with a CSV file of population data, located on GitHub. The final R dataframe will have the following columns:

Title
gender_person
Rating



```{r,warning=FALSE,message=FALSE}
library(RMySQL)
library(tidyverse)
library(dplyr)
library(DBI)
```


Getting and Preparing the Data

Step 1. Connect to MySQL and retrieve the tb dataset stored in a database table.

```{r movies-connect,eval=TRUE}
mydb = dbConnect(MySQL(), user='root', password='Albania777', dbname='movies_db', host='localhost')
```

Return the movies query below and store the results a dataframe called movies

```{r}
movies.df <- dbGetQuery(mydb, "select title, gender_person,rating from movies_observ")
```


```{r}
names(movies.df) 
```


```{r movies-summary,eval=TRUE}
summary(movies.df)
```

```{r movies-sum,eval=TRUE}
print(movies.df)
```





```{r}
qplot(title, rating, data=movies.df,xlab = "Rating", ylab = "Movie", main = "Individual Movie Rating by Gender") + facet_wrap(~gender_person) + theme(axis.text.x = element_text(angle = 90, hjust = 1))
```





```{r}
ggplot(movies.df, aes(x = reorder(title, rating), y = rating, fill = title), xlab = 'Rating',  col = I("grey")) + geom_bar(stat = "identity") + 
  ggtitle("Movie Cummulative Ratings") +  labs(x = "Movie") +  coord_flip()

```

```{r db-disconnect,eval=TRUE}
dbDisconnect(mydb)
```