---
title: "Project 1"
author: "Erinda Budo"
date: "9/18/2019"
output:
  html_document: default
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

In this project, we’re given a text file with chess tournament results where the information has some structure.The task is to generate a .CSV file (that could for example be imported into a SQL database) with the following information for all of the players:

 PlayerName  | State |TotalPoints|PreRating |AvgOppPreRating |
-------------| ------|-----------|----------|----------------|
 Gary Hua    | ON    |  6.0      | 1794     |  1605          |



------------------------------------------------------------------------------------------

```{r,warning=FALSE,message=FALSE} 
#Load necessary libraries
library(stringr)
library(DT)
library(dplyr)
library(tidyverse)
```


**LOAD DATA**

```{r}
# Load the data from the file in my github repository
get_url <- 'https://raw.githubusercontent.com/ErindaB/Data607_Assignment/master/tournamentinfo.txt'
raw_data <- read.delim(get_url, header=FALSE, stringsAsFactors =FALSE )
head(raw_data)
```


The chess tournament text file is written in a format that is not legible in R as it is. This file needs to be restructured in order to find the average pre-tournament score.


```{r}
#Remove the header which is comprised of two rows
raw_data <- raw_data[c(5:nrow(raw_data)),]
head(raw_data)
```

**Extracting main variables from the file**

```{r}
id_var <- "\\d{1,2}(?=\\s\\|)"
chess_id <- unlist(str_extract_all(unlist(raw_data), id_var))
```


*Extract player's names*
```{r}
# Extract the player's names 
# Get names in  capital letters, followed by a space and with at least two matches

name <- "([[:upper:]]+\\s){2,}"
player_name <- unlist(str_extract_all(unlist(raw_data),name))
str_trim(player_name, side = "right")
```



*Extract player's total points*

```{r}
# Extract each player's total number of points

points <- "\\d\\.\\d"
total_points <- unlist(str_extract_all(unlist(raw_data), points))
total_points
```

*Extract states*

```{r}
# Extract each player's state
# Get state by  2 capital letters,followed by a space and  '|'

state <- "([[:upper:]]){2}\\s(?=\\|)"
states <- unlist(str_extract_all(unlist(raw_data), state))
str_trim(states, side = "right")
```


*Extract pre-rating*
```{r}
#  Extract each player's pre-rating
#  Avoid  patterns with a > and a space. Then search for between 1 and two spaces, or one space and a colon, followed by 3 or 4 digits, and  a space or the letter 'P'

feature <- "(?<!\\>\\s)(?<=\\s{1,2}|\\s\\:)(\\d{3,4}(?=\\s|P))"
pre_rating <- unlist(str_extract_all(unlist(raw_data), feature))
pre_rating <- str_trim(pre_rating)
pre_rating
```


*Extract average of opponents pre-rating*
```{r}

opponents_pattern <- "(\\d{1,}|[[:blank:]]{1})(?=\\|)"
player_opponents <- unlist(str_extract_all(unlist(raw_data), opponents_pattern))
player_opponents[player_opponents==" "]  <- NA
opponent1 <- as.numeric(player_opponents[seq(4, length(player_opponents), 10)])
opponent1 <- as.numeric(opponent1[seq(1, length(opponent1), 2)])
opponent2 <- as.numeric(player_opponents[seq(5, length(player_opponents), 10)])
opponent2 <- as.numeric(opponent2[seq(1, length(opponent2), 2)])
opponent3 <- as.numeric(player_opponents[seq(6, length(player_opponents), 10)])
opponent3 <- as.numeric(opponent3[seq(1, length(opponent3), 2)])
opponent4 <- as.numeric(player_opponents[seq(7, length(player_opponents), 10)])
opponent4 <- as.numeric(opponent4[seq(1, length(opponent4), 2)])
opponent5 <- as.numeric(player_opponents[seq(8, length(player_opponents), 10)])
opponent5 <- as.numeric(opponent5[seq(1, length(opponent5), 2)])
opponent6 <- as.numeric(player_opponents[seq(9, length(player_opponents), 10)])
opponent6 <- as.numeric(opponent6[seq(1, length(opponent6), 2)])
opponent7 <- as.numeric(player_opponents[seq(10, length(player_opponents), 10)])
opponent7 <- as.numeric(opponent7[seq(1, length(opponent7), 2)])
player_opponents <- matrix(c(opponent1, opponent2, opponent3, opponent4, opponent5, opponent6, opponent7),nrow = 64, ncol = 7)

# Match the opponent to the player's id

AvgOppRating<- 0
chess_tournament <- 0
for (i in 1:(length(chess_id)))
{
  AvgOppRating[i] <- mean(as.numeric(pre_rating[player_opponents[i,]]), na.rm = T)
}


```


*Create a table with all new variables*

```{r}
AvgOppRating<-round(AvgOppRating)
chess_tournament <- data.frame(player_name,states, total_points, pre_rating, AvgOppRating)
chess_tournament
```

**Write info to .csv file**

```{r}
write.csv(chess_tournament, file = "chess_tournament.csv")

```

**Exploratory Plot**

```{r}
p <- ggplot(chess_tournament, aes(pre_rating, AvgOppRating)) + geom_point(aes(color=total_points)) + ggtitle("Comparing Player Pre-Rating to Avg Opponent Pre-Rating \n  by Total Points Gained")
p

```


